---
title: "Appendix"
author: "Eric Marcon, Florence Puech"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    theme: sandstone
    toc: yes
    toc_float: yes
---
  
```{r Options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, echo = TRUE, tidy=FALSE, tidy.opts=list(blank=FALSE, width.cutoff=80), out.width='\\maxwidth')
options(width=80)
# Install necessary packages
CRANLibrary <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Install Packages from GitHub
GitHubLibrary <- function(Packages) {
  InstallAndLoad <- function(Package) {
    Package_split <- stringr::str_split(Package, "/", simplify = TRUE)
    if (!Package_split[1, 2] %in% installed.packages()[, 1]) {remotes::install_github(Package)}
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Necessary packages listed here
CRANLibrary(c("dbmss", "entropart", "formatR", "kableExtra", "maptools", "rgdal", "rgeos", "tidyverse"))
GitHubLibrary("EricMarcon/SpatDiv")
```

This document contains the code used to produce the figures of the paper and run all examples.
It is part of an RStudio project with R Markdown documents.


# Preparing data

## Download data

Data are available on https://opendata.paris.fr/explore/dataset/les-arbres/export/

Preparing data consists of downloading open-data files and wrangling it to obtain:

- `the_trees`: a tibble with one line per tree, described mainly by its species, coordinates, and the green space it belongs to.
- `Paris`: an `owin` object, i.e. a geometric window with the limits of the city.

Those files are cached in the `data` folder.
The wrangling process can be reinitialized by deleting it, e.g. to benefit from an update of the original data files.
  
The GeoJSON dataset that contains the trees is downloaded to `/data/les-arbres.les-arbres.geojson`.

```{r}
# Create the data folder if necessary
if (!dir.exists("data")) dir.create("data")
# Download the dataset if necessary (100)
if (!file.exists("data/ParisTrees.RData"))
  download.file(
    "https://opendata.paris.fr/explore/dataset/les-arbres/download/?format=geojson", 
    destfile="data/les-arbres.geojson")
```

Another dataset contains the limits of arrondissements, i.e. Paris administrative sectors.
```{r}
# Arrondissements
if (!file.exists("data/ParisTrees.RData"))
  download.file(
    "https://opendata.arcgis.com/datasets/8463ae61c7ba4388bf08189effcbf2f1_0.geojson",
    destfile="data/arrondissements.geojson")
```


## Datum

GeoJSON datasets are read by the **rgdal** package.

```{r, message=FALSE}
if (!file.exists("data/ParisTrees.RData")) {
  library("rgdal")
  the_trees_GeoJSON <- readOGR("data/les-arbres.geojson", 
                             layer = "les-arbres", 
                             encoding = "UTF-8", use_iconv = TRUE)
  arrondissements <- readOGR("data/arrondissements.geojson", 
                           layer="COMMUNE_ARRONDISSEMENT_BOIS", 
                           encoding = "UTF-8", use_iconv = TRUE)
}
```

Data are projected into Lambert 93 coordinates.

```{r, warning=FALSE, message=FALSE}
if (!file.exists("data/ParisTrees.RData")) {
  library("magrittr")
  the_trees_GeoJSON %<>% 
    spTransform(CRS("+init=epsg:2154"))
  arrondissements %<>% 
    spTransform(CRS("+init=epsg:2154"))
}
```

Latitude and longitude are now in meters.


## The trees

Useful data are gathered in a tibble to simplify further operations.
Column names are translated into English. 
So are green-space types. 

```{r, tidy=FALSE, message=FALSE}
if (!file.exists("data/ParisTrees.RData")) {
  library("dplyr")
  library("stringr")
  library("forcats")
  the_trees_GeoJSON@data %>%
    as_tibble() %>% 
    # Bind Lambert coordinates
    bind_cols(as.data.frame(the_trees_GeoJSON@coords)) %>% 
    # Delete useless columns
    select(-idemplacement, -typeemplacement, -geo_point_2d1, -geo_point_2d2, 
           -complementadresse, -stadedeveloppement) %>% 
    # Retain intra-muros Paris data only
    dplyr::filter(str_detect(arrondissement, "^PARIS")) %>% 
    # Add a full species name column
    mutate(SpeciesName=as.factor(paste(genre, espece))) %>% 
    # Factorize some columns
    mutate(libellefrancais=as.factor(libellefrancais)) %>% 
    mutate(genre=as.factor(genre)) %>% 
    mutate(espece=as.factor(espece)) %>% 
    mutate(remarquable=(remarquable=="OUI")) %>% 
    # Translate green space types
    mutate(domanialite=as.factor(domanialite)) %>% 
    mutate(domanialite=fct_recode(domanialite, 
             "Street" = "Alignement", 
             "Park" = "Jardin", 
             "Cemetery" = "CIMETIERE", 
             "School" = "DASCO", 
             "Ring Motorway" = "PERIPHERIQUE", 
             "Sports facility" = "DJS", 
             "Nursery" = "DFPE", 
             "Cultural facility" = "DAC", 
             "Social Services" = "DASES")) %>% 
    # Arrondissement reduced to a number
    mutate(arrondissement=str_replace(arrondissement, "PARIS ", "")) %>% 
    mutate(arrondissement=str_replace(arrondissement, "E ARRDT", "")) %>% 
    mutate(arrondissement=str_replace(arrondissement, "ER ARRDT", "")) %>%  
    mutate(arrondissement=as.integer(arrondissement)) %>% 
    # Simplify cemetery addresses to have a single one per cemetery
    mutate(adresse=str_trim(ifelse(str_detect(as.character(adresse), "^CIMETIERE "), 
      str_split(as.character(adresse), "/", simplify=TRUE)[, 1], 
      as.character(adresse)))) %>% 
    # Avoid duplicate addresses by adding arrondissement and domanialite data to them
    mutate(adresse = as.factor(paste(adresse, arrondissement, domanialite))) %>% 
    # Rename columns
    rename(ID=idbase, 
           FrenchSpeciesName=libellefrancais,
           GreenSpaceType=domanialite,
           Address=adresse,
           Arrondissement=arrondissement,
           Circumference=circonferenceencm,
           Height=hauteurenm,
           Remarkable=remarquable,
           Genus=genre,
           Species=espece,
           x=coords.x1,
           y=coords.x2,
           VarietyOrCultivar=varieteoucultivar) %>% 
    # Reorder them
    select(ID, SpeciesName, FrenchSpeciesName, GreenSpaceType, Address, Arrondissement,
           Circumference, Height, Remarkable, Genus, Species, x, y,
           VarietyOrCultivar) ->
    the_trees
}
```


## Arrondissements

```{r}
if (!file.exists("data/ParisTrees.RData")) {
  # Retain intra-muros Paris data only (postal code = 75xxx). Eliminate outer forests ("Bois")
  arrondissements <- arrondissements[arrondissements@data$C_CAINSEE %/% 1000 == 75 & 
                                       arrondissements@data$B_BOIS == "N" , ]
  arrondissements <- arrondissements[arrondissements@data$C_CAINSEE %/% 1000 == 75 & 
                                       arrondissements@data$B_BOIS == "N" , ]
  # Reduce arrondissement name to its number
  arrondissements@data$L_CAB %<>% 
    str_replace("er", "") %>% 
    str_replace("e", "") %>% 
    str_replace("ème", "") %>% 
    as.integer()
  # Dissolve arrondissement borders to get a single polygon for Paris
  library("maptools")
  Paris <- unionSpatialPolygons(arrondissements, 
                                rep(TRUE, length(arrondissements@polygons)))
}
```

## Save the data

```{r}
if (!file.exists("data/ParisTrees.RData"))
  save(the_trees, Paris, file = "data/ParisTrees.RData")
```


# Exploratory analyzes

```{r}
# Load the data
load("data/ParisTrees.RData")
```

## Species abundances

The total number of trees is `r nrow(the_trees)`.

The most abundant species are:

```{r}
library("dplyr")
the_trees %>% 
  group_by(SpeciesName) %>% 
  summarise(Number=n(), .groups='drop') %>% 
  arrange(desc(Number))%>% 
  knitr::kable(caption="Domanialité", longtable = FALSE, booktabs = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)
```



# Spatial concentration

## Weighted, Marked, Planar Point Pattern

```{r}
# Prepare the window
library("maptools")
library("dbmss")
Paris_owin <- as.owin(Paris)
# Make a wmppp object
the_trees %>% 
  select(x, y, SpeciesName) %>% 
  rename(PointType = SpeciesName) %>% 
  as.wmppp(window = Paris_owin) ->
  the_trees_wmppp
# Plot the map without the too-many trees (alpha=0)
autoplot(the_trees_wmppp, alpha = 0) +
  # Add the trees as small black points
  geom_point(aes(x = x, y = y), size = 0.1, stroke = 0) +
  # Hide the legend
  guides(color = FALSE, size = FALSE)
```

## Concentration

M function applied to the whole dataset.

```{r}
M_Platanus <- Mhat(the_trees_wmppp, ReferenceType = "Platanus x hispanica")
autoplot(M_Platanus)
```


# Diversity

In entropart, data are organized in metacommunity object where each line is a species and each column a community (here, a street).

```{r}
library("tidyr")
library("entropart")
the_trees %>% 
  # Retain the streets
  dplyr::filter(GreenSpaceType == "Street") %>% 
  select(SpeciesName, Address, ID)  %>% 
  pivot_wider(names_from = Address, 
            values_from = ID,
            values_fill = 0,
            values_fn = length) %>% 
  MetaCommunity(Weights = colSums(.[, -1])) %>% 
  # Suppress warnings due to sample coverage estimation
  suppressWarnings -> streets_MC
```


Diversity of the streets :
```{r}
DivPart(q = 1, streets_MC) -> streets_diversity
streets_diversity$TotalAlphaDiversity
streets_diversity$TotalBetaDiversity
streets_diversity$GammaDiversity
```


# Spatially explicit diversity

## Make a point pattern (wmppp) from its address

```{r, tidy=FALSE, message=FALSE}
#' wmppp from an address
#'
#' @param green_spaces A dataframe with address, x and y columns
#' @param address An address to be found in `green_spaces$address`
#' @param unitname A vector with the name of the unit of distance (singular, plural)
#'
#' @return A `wmppp` object
wmppp_address <- function(green_spaces, address, unitname = c("meter", "meters")) {
  # Prepare the dataframe
  green_spaces %>%
    dplyr::filter(Address == address) %>%
    dplyr::select(x, y,
           PointType=SpeciesName,
           PointWeight=Circumference) %>%
    dbmss::wmppp(unitname = unitname)
}
```


## Parc des Buttes Chaumont

```{r}
Buttes_Chaumont <- wmppp_address(the_trees, "PARC DES BUTTES CHAUMONT 19 Park")
library("SpatDiv")
Buttes_Chaumont$window <- alphahull(Buttes_Chaumont)
autoplot(Buttes_Chaumont, labelSize = "Circumference", labelColor = "Species")
```


## Local diversity

```{r}
Buttes_Chaumont_accum <- Mixing(Buttes_Chaumont, q.seq=c(0,1), n.seq = c(5, 10),  Individual=TRUE)
plot(Buttes_Chaumont_accum, q=1)
MapPlot(Buttes_Chaumont_accum, Order=1, NeighborHood=11)
```

