---
title: "Appendix"
author: "Eric Marcon, Florence Puech"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    theme: sandstone
    toc: yes
    toc_float: yes
---
  
```{r Options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, echo = TRUE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=80), out.width='\\maxwidth')
options(width=80)
# Install necessary packages
CRANLibrary <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
    #require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Install Packages from GitHub
GitHubLibrary <- function(Packages) {
  InstallAndLoad <- function(Package) {
    Package_split <- stringr::str_split(Package, "/", simplify = TRUE)
    if (!Package_split[1, 2] %in% installed.packages()[, 1]) {remotes::install_github(Package)}
    #require(Package_split[1, 2], character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Necessary packages listed here
CRANLibrary(c("entropart", "formatR", "kableExtra", "maptools", "rgdal", "rgeos", "tidyverse"))
GitHubLibrary("EricMarcon/SpatDiv")
```

This document contains the code used to produce the figures of the paper and run all examples.
It is part of an RStudio project with R Markdown documents.


# Preparing data

## Download data

Data are available on https://opendata.paris.fr/explore/dataset/les-arbres/export/
  
The GeoJSON dataset that contains the trees is downloaded to `/data/les-arbres.les-arbres.geojson`.

```{r}
# Create the data folder if necessary
if (!dir.exists("data")) dir.create("data")
# Download the dataset if necessary (100)
if (!file.exists("data/les-arbres.geojson"))
  download.file("https://opendata.paris.fr/explore/dataset/les-arbres/download/?format=geojson", destfile="data/les-arbres.geojson")
```

Another dataset contains the limits of arrondissements, i.e. Paris administrative sectors.
```{r}
# Arrondissements
if (!file.exists("data/arrondissements.geojson"))
  download.file("https://opendata.arcgis.com/datasets/8463ae61c7ba4388bf08189effcbf2f1_0.geojson", destfile="data/arrondissements.geojson")
```


## Datum

GeoJSON datasets are read by the **rgdal** package.

```{r}
library("rgdal")
the_trees_GeoJSON <- readOGR("data/les-arbres.geojson", layer = "les-arbres", encoding = "UTF-8", use_iconv = TRUE)
arrondissements <- readOGR("data/arrondissements.geojson", layer="COMMUNE_ARRONDISSEMENT_BOIS", encoding = "UTF-8", use_iconv = TRUE)
```

Data are projected into Lambert 93 coordinates.

```{r, warning=FALSE}
library("magrittr")
the_trees_GeoJSON %<>% 
  spTransform(CRS("+init=epsg:2154"))
arrondissements %<>% 
  spTransform(CRS("+init=epsg:2154"))
```

Latitude and longitude are now in meters.

## The trees

Useful data are gathered in a tibble to simplify further operations.
Column names are translated into English. 
So are green space types. 

```{r, tidy=FALSE}
library("tidyverse")
the_trees_GeoJSON@data %>%
  as_tibble() %>% 
  # Bind Lambert coordinates
  bind_cols(as.data.frame(the_trees_GeoJSON@coords)) %>% 
  # Delete useless columns
  select(-idemplacement, -typeemplacement, -geo_point_2d1, -geo_point_2d2, -complementadresse, -stadedeveloppement) %>% 
  # Retain intra-muros Paris data only
  dplyr::filter(str_detect(arrondissement, "^PARIS")) %>% 
  # Add a full species name column
  mutate(SpeciesName=as.factor(paste(genre, espece))) %>% 
  # Factorize some columns
  mutate(libellefrancais=as.factor(libellefrancais)) %>% 
  mutate(genre=as.factor(genre)) %>% 
  mutate(espece=as.factor(espece)) %>% 
  mutate(remarquable=(remarquable=="OUI")) %>% 
  # Translate green space types
  mutate(domanialite=as.factor(domanialite)) %>% 
  mutate(domanialite=fct_recode(domanialite, 
           "Street" = "Alignement", 
           "Park" = "Jardin", 
           "Cemetery" = "CIMETIERE", 
           "School" = "DASCO", 
           "Ring Motorway" = "PERIPHERIQUE", 
           "Sports facility" = "DJS", 
           "Nursery" = "DFPE", 
           "Cultural facility" = "DAC", 
           "Social Services" = "DASES")) %>% 
  # Arrondissement reduced to a number
  mutate(arrondissement=str_replace(arrondissement, "PARIS ", "")) %>% 
  mutate(arrondissement=str_replace(arrondissement, "E ARRDT", "")) %>% 
  mutate(arrondissement=str_replace(arrondissement, "ER ARRDT", "")) %>%  
  mutate(arrondissement=as.integer(arrondissement)) %>% 
  # Simplify cemetery addresses to have a single one per cemetery
  mutate(adresse=str_trim(ifelse(str_detect(as.character(adresse), "^CIMETIERE "), str_split(as.character(adresse), "/", simplify=TRUE)[, 1], as.character(adresse)))) %>% 
  # Avoid duplicate addresses by adding arrondissement and domanialite data to them
  mutate(adresse = as.factor(paste(adresse, arrondissement, domanialite))) %>% 
  # Rename columns
  rename(ID=idbase, 
         FrenchSpeciesName=libellefrancais,
         GreenSpaceType=domanialite,
         Address=adresse,
         Arrondissement=arrondissement,
         Circumference=circonferenceencm,
         Height=hauteurenm,
         Remarkable=remarquable,
         Genus=genre,
         Species=espece,
         x=coords.x1,
         y=coords.x2,
         VarietyOrCultivar=varieteoucultivar) ->
  the_trees
```

## Arrondissements

```{r}
# Retain intra-muros Paris data only (postal code = 75xxx). Elmiminate outer forests ("Bois")
arrondissements <- arrondissements[arrondissements@data$C_CAINSEE %/% 1000 == 75 & arrondissements@data$B_BOIS == "N" , ]
arrondissements <- arrondissements[arrondissements@data$C_CAINSEE %/% 1000 == 75 & arrondissements@data$B_BOIS == "N" , ]
# Reduce arrondissement name to its number
arrondissements@data$L_CAB %<>% 
  str_replace("er", "") %>% 
  str_replace("e", "") %>% 
  str_replace("ème", "") %>% 
  as.integer()
# Dissolve arrondissement borders to get a single polygon for Paris
library("maptools")
Paris <- unionSpatialPolygons(arrondissements, rep(TRUE, length(arrondissements@polygons)))
```




# Exploratory analyzes

## Species abundances

The total number of trees is `r nrow(the_trees)`.

The most abundant species are

```{r}
the_trees %>% 
  group_by(SpeciesName) %>% 
  summarise(Number=n(), .groups='drop') %>% 
  arrange(desc(Number)) %>% 
  print ->
  AbdFreqCount
```



# Spatial concentration

## Weighted, Marked, Planar Point Pattern

```{r}
# Prepare the window
library("dbmss")
Paris_owin <- as.owin(Paris)
# Make a wmppp object
the_trees %>% 
  select(x, y, SpeciesName) %>% 
  rename(PointType=SpeciesName) %>% 
  as.wmppp(window =Paris_owin) -> 
  the_trees_wmppp
autoplot(the_trees_wmppp, palette="Set1", LegendLabels = c("Circumference", "Species"), alpha = 0) +
  geom_point(aes(x=x, y=y, size=0.1, stroke=0)) +
  guides(color = FALSE, size = FALSE)
```
## Concentration

```{r}
M_Platanus <- Mhat(the_trees_wmppp, ReferenceType = "Platanus x hispanica")
autoplot(M_Platanus)
```

